---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r load_data, echo=TRUE}
adata=read.csv(unzip("./activity.zip"),header=TRUE, sep=",")
adata$date=as.Date(adata$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

1. Total number of steps taken per day.

```{r total_steps, echo=TRUE}
library(dplyr)
day_group<-group_by(adata, date)
step_totals<-summarize(day_group, total=sum(steps))
step_totals
```

2. Histogram of total number of steps taken per day.
```{r histogram_total, echo=TRUE }
library(ggplot2)
ggplot(step_totals, aes(x=total)) + geom_histogram(binwidth = 3000)
```

3. Calulae the mean and median of total number of steps taken per day.
```{r mean_and_median, echo=TRUE}
mean_totalsteps<-mean(step_totals$total,na.rm=TRUE)
mean_totalsteps
median_totalsteps<-median(step_totals$total,na.rm=TRUE)
median_totalsteps
```

## What is the average daily activity pattern?

1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
 Note 2012 was a leap year with 366 days = 527040 minutes / 5 = 105408
 October 1st is Julian date 275 = 396000 minutes / 5 = 79200

```{r time_series, echo=TRUE}

```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
